---
title: "With data from API"
author: "Alejandra Gerosa"
date: "3/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(jsonlite)
library(glue)
library(skimr)

```

## Pedcount

Let's import the newest dataset and apply the tidying steps we found before.

The longitude and latitude need to be extracted differently, because the_geom is a dataframe this time.

```{r read}
pedcount <- fromJSON("https://data.cityofnewyork.us/resource/cqsj-cfgu.json")
pedcount <- pedcount %>% mutate(
  lon = map(the_geom$coordinates, 1),
  lat = map(the_geom$coordinates, 2))


```

Now let's make it tidy

Note: setting up the columns in the second gather call here requires fiddling each time the file is updated. It'd be ideal to find a fix

Also, they got rid of all the date columns!!

```{r tidy, echo=FALSE}

pedcount_tidy <- pedcount %>%
  select(lon, lat, everything(), -the_geom) %>%
  gather(key = "measurement_coding", value = "ped_count", 10:83) %>% 
  mutate(
    ped_count = as.numeric(ped_count),
    type_of_measurement = case_when(
      str_sub(measurement_coding, -1, -1) == "d" ~ "Weekend Noon",
      str_sub(measurement_coding, -2, -1) == "am" ~ "Weekday Morning",
      str_sub(measurement_coding, -2, -1) == "pm" ~ "Weekday Evening"),
    season = case_when(
      str_sub(measurement_coding, 1, 3) == "may" ~ "Spring",
      str_sub(measurement_coding, 1, 3) == "sep" ~ "Fall"),
    year = as.numeric(
      str_sub(measurement_coding, -5, -4)
      )
    + 2000,
    year_month = case_when(
      season == "Spring" ~ glue("{year}/05"),
      season == "Fall" ~ glue("{year}/09")
    )
    )

```

"NAs introduced by coercion". Let's find out why and if there were actual numbers there that we need to fix.
```{r fix_NAs1}
pedcount_tidy %>% count(year_month)
pedcount_tidy %>% count(borough)

```
So, we know that there are 342 observations for all months other than September 2018.

```{r fix_NAs2}
pedcount_tidy %>% filter(is.na(ped_count)) %>% count(year_month)

```

Let's take a look at observations from one of the times with a ton of "missing" data and from one that has mostly good data to see what was there originally. Say, September 2007 vs September 2013.

```{r fix_NAs3}
pedcount %>% select(starts_with("sept_07"), starts_with("sept_13"))

```
So now we know that commas are a major problem that prevented as.numeric() from reading in a lot of these. Let's redo the tidy step and then check out the NAs again.

```{r tidy2}
pedcount_tidy <- pedcount %>%
  select(lon, lat, everything(), -the_geom) %>%
  gather(key = "measurement_coding", value = "ped_count", 10:83) %>% 
  mutate(
    ped_count = as.numeric(str_replace(ped_count, ",", "")),
    type_of_measurement = case_when(
      str_sub(measurement_coding, -1, -1) == "d" ~ "Weekend Noon",
      str_sub(measurement_coding, -2, -1) == "am" ~ "Weekday Morning",
      str_sub(measurement_coding, -2, -1) == "pm" ~ "Weekday Evening"),
    season = case_when(
      str_sub(measurement_coding, 1, 3) == "may" ~ "Spring",
      str_sub(measurement_coding, 1, 3) == "sep" ~ "Fall"),
    year = as.numeric(
      str_sub(measurement_coding, -5, -4)
      )
    + 2000,
    year_month = case_when(
      season == "Spring" ~ glue("{year}/05"),
      season == "Fall" ~ glue("{year}/09")
    )
    )

pedcount_tidy %>% filter(is.na(ped_count)) %>% count(year_month)

```
This looks way more reasonable and like they are probably actually missing, except for May 2019. Let's check that and another one just in case.

```{r}
pedcount %>% select(starts_with("sept_08"), starts_with("may_19"))

```
Yes, these are actually missing.

 Also, lon and lat are still lists, I should probably fix that.

# Analyze

```{r boxplots}

pedcount_tidy %>% ggplot(
  aes(x = year_month, y = log(ped_count))
) +
  geom_boxplot() +
  facet_wrap("season", ncol = 1) +
  theme_minimal()

```
